A sustainable approach to tourist signage on heritage trails

### 2023-11-15 paper.tex
Pre trasporto di QR-code

% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
\address{%
$^{1}$ \quad University of Pisa (retired); augusto.ciuffoletti@unipi.it\\
$^{2}$ \quad University of Pisa; mariagrazia.deri@sp.unipi.it\\
$^{3}$ \quad University of Pisa; letizia.chiti@}

@@

% Abstract (Do not insert blank lines, i.e. \\)
\abstract{Understanding the cultural aspects of an area rich in heritage is crucial to building a lasting educational experience from an excursion. Many articles in literature explore the use of sophisticated technologies to achieve such a goal. Tourism proposals for inland areas are significant because the presence of human artifacts and signage can harm the experience and create pollutants. Through a holistic methodology and analysis, this paper examines the signage for an area rich in cultural and natural assets: the study encompasses history, touristic vocation, and the environmental context. According to the analysis, slow, community-involved tourism is the preferred destination, and the signage solution must meet strict sustainability requirements in the social, economic, and environmental realms. After applying the appropriate governance guidelines provided, the QR code technology is selected for a thoroughly documented experimental deployment.}
%\abstract{A single paragraph of about 200 words maximum. For research articles, abstracts should give a pertinent overview of the work. We strongly encourage authors to use the following style of structured abstracts, but without headings: (1) Background: place the question addressed in a broad context and highlight the purpose of the study; (2) Methods: describe briefly the main methods or treatments applied; (3) Results: summarize the article's main findings; (4) Conclusions: indicate the main conclusions or interpretations. The abstract should be an objective representation of the article, it must not contain results that are not presented and substantiated in the main text and should not exaggerate the main conclusions.}

% Keywords
\keyword{Inland areas; heritage trail; digital sustainability; slow tourism; destination governance; community-involved tourism; cultural heritage; holistic approach; QR-code; trail signage}

@@

Several promotional websites, like \citeWeb{tbdl} and \citeWeb{valdilima} advertise \emph{Bagni di Lucca} attractions. The sitography closing this paper contains a longer yet incomplete list of touristic initiatives  \citeWeb{agu,r2o,ronda,garfraft,valdilima,canyonpark,experience,ebikea,ebikel,bdlt2,bdl}.

@@

We aim to enhance the tourist experience in an inland area located around the village of \emph{Casoli}. Our goal is tightly related with the fourth pillar of sustainability: there are natural and cultural assets that need to be nurtured or at least preserved. We concentrate on a signage system, the realization of which is finalized to reach such a cultural result.

To avoid the over-tourism pitfall, we need to take into account the other three pillars: social, environmental, and economic. To this end, we analyze the social fabric that will host our initiative, the environment where signage will be deployed, and the economic framework that will support its implementation. From this study, we derive the requirements for our signage system: low environmental impact, open to participation of the residents, responsive to the visitor's information needs, and economically sustainable with a minimal tourist flow.

The approach we applied to reach such conclusions merely consists of designing the system with sustainability in mind from the early stages. The simplicity of the description must not mislead: its implementation is challenging, requiring the collaboration among experts in distinct disciplines converging to a final result.

@@

An outcome of the analysis of the social fabric is that, in order to preserve the native social and economic framework, mass tourism is not a candidate target, while a motivated, non-casual visitor is preferable. Regarding the potential reasons for a visit, we found a cave, reachable with some effort, inhabited since pre-history, a village with historical buildings, and nearby architecture dating back to the Middle Ages. Such targets offer a variety of access ways, from those requiring performance to relaxing ones. For this reason, we opted not to suggest a geo-itinerary, allowing the user cherry-picking locations based on personal interests. In this way we expect, legitimated by the results of \cite{sol18a}, that the educational impact of the signage is maximized. Furthermore, our choice aims at concise and effective communication with the tourist  by inserting within the QR code a text of an average of 460 characters that facilitates readability and immediately provides the information without further steps to web pages, which may slow down the visit or discourage the less motivated visitor.

@@

%\nociteWeb{*}
\bibliographystyleWeb{abbrv}
\bibliographyWeb{biblio/sitography.bib}

>>>>

% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
\address{%
$^{1}$ \quad University of Pisa (retired); augusto.ciuffoletti@unipi.it\\
$^{2}$ \quad University of Pisa; mariagrazia.deri@sp.unipi.it\\
$^{3}$ \quad University of Pisa; letizia.chiti@cfs.unipi.it}

@@

% Abstract (Do not insert blank lines, i.e. \\)
\abstract{Understanding the cultural aspects of an area rich in heritage is crucial to building a lasting educational experience from an excursion. Many articles in the literature explore the use of sophisticated technologies to achieve such a goal. Tourism proposals for inland areas are significant because the presence of human artifacts and signage can harm the experience and create pollutants. Through a holistic methodology and analysis, this paper examines the signage for an area rich in cultural and natural assets: the study encompasses history, touristic vocation, and the environmental context. According to the analysis, slow, community-involved tourism is the preferred destination, and the signage solution must meet strict sustainability requirements in the social, economic, and environmental realms. After applying the appropriate governance guidelines provided, the QR code technology is selected for a thoroughly documented experimental deployment.}
%\abstract{A single paragraph of about 200 words maximum. For research articles, abstracts should give a pertinent overview of the work. We strongly encourage authors to use the following style of structured abstracts, but without headings: (1) Background: place the question addressed in a broad context and highlight the purpose of the study; (2) Methods: describe briefly the main methods or treatments applied; (3) Results: summarize the article's main findings; (4) Conclusions: indicate the main conclusions or interpretations. The abstract should be an objective representation of the article, it must not contain results that are not presented and substantiated in the main text and should not exaggerate the main conclusions.}

% Keywords
\keyword{Inland areas; digital sustainability; slow tourism; destination governance; community-involved tourism; cultural heritage; holistic approach; QR-code}

@@

Several promotional websites, like \citeWeb{tbdl} and \citeWeb{valdilima} advertise \emph{Bagni di Lucca} attractions. The sitography closing this paper contains a longer yet incomplete list of touristic initiatives  \citeWeb{valdilima,agu,r2o,ronda,garfraft,canyonpark,experience,ebikea,ebikel,bdlt2,bdl}.

@@

We aim to enhance the tourist experience in an inland area located around the village of \emph{Casoli}. Our goal is closely related to the fourth pillar of sustainability: there are cultural assets that need to be nurtured or at least preserved. The proposed signage system aims to achieve such a goal by promoting the documentation and touristic fruition of these assets.

To avoid the over-tourism pitfall, we need to take into account the other three pillars: social, environmental, and economic. To this end, we analyze the social fabric that will host our initiative, the environment where signage will be deployed, and the economic framework that will support its implementation. From this study, we derive the requirements for our signage system: low environmental impact, open to the participation of the residents, responsive to the visitor's information needs, and economically sustainable with minimal tourist flow.

The approach we applied to reach such conclusions merely consists of designing the system with sustainability in mind from the early stages. The simplicity of the description must not mislead: its implementation is challenging, requiring collaboration among experts in distinct disciplines to converge to a final result.

@@

An outcome of the analysis of the social fabric is that, to preserve the native social and economic framework, mass tourism is not a candidate target, while a motivated, non-casual visitor is preferable. Regarding the potential reasons for a visit, we found a cave, reachable with some effort, inhabited since pre-history, a village with historical buildings, and nearby architecture dating back to the Middle Ages. Such targets offer a variety of access ways, from those requiring performance to relaxing ones. For this reason, we opted not to suggest a geo-itinerary, allowing the user to cherry-pick locations based on personal interests. In this way we expect, legitimated by the results of \cite{sol18a}, that the educational impact of the signage is maximized. Furthermore, our choice aims at concise and effective communication with the tourist by inserting within the QR code a text of an average of 460 characters that facilitates readability and immediately provides the information without further steps to web pages, which may slow down the visit or discourage the less motivated visitor.

@@

%\nociteWeb{*}
\bibliographystyleWeb{unsrt}
\bibliographyWeb{biblio/sitography.bib}
